---
title: "Assignment Lesson 7"
author: "Team Zeven: Rob Maas & Wilmar van Ommeren"
date: "Wednesday, January 14, 2015"
output:
  html_document:
    theme: cosmo
    toc: yes
    toc_depth: 4
    number_sections: true
---

```{r table}
```{r, echo=FALSE, message=FALSE}
library(knitr)
opts_chunk$set(fig.width=7, fig.align='center', fig.height=7, dpi=100)
```
# Description
This Rpub Document includes the assignment of [Geo-Scripting Lesson 7](https://geoscripting-wur.github.io/AdvancedRasterAnalysis/). In this Lesson LandSat bands are used to predict the tree cover with a linear regression model.

Link to github: [https://github.com/Boonalds/Lesson7_TeamZeven.git](https://github.com/Boonalds/Lesson7_TeamZeven.git)

# Methodology
## Plot the relationship between the LandSat bands and the VCF tree cover
For the first part of this exercise we use two libraries, some LandSat
bands and the VCF Tree Cover product.

```{r loading, message=FALSE, include=TRUE, results='hide'}
# Load libraries
library(sp) 
library(raster)
library(RColorBrewer)

# Load data
load("Data/GewataB1.rda")
load("Data/GewataB5.rda")
load("Data/GewataB7.rda")
load("Data/vcfGewata.rda")
```

```{r loading extra, message=FALSE, include=FALSE, results='hide'}
load("Data/GewataB2.rda")
load("Data/GewataB3.rda")
load("Data/GewataB4.rda")
load("Data/trainingPoly.rda")
```

Not all the data in the VCF product is correct, some pixels contain more then 100 percent of tree cover(as can be seen in the histogram). This erroneous data is probably  caused by cloudcover, water or cloudshadow pixels. Before plotting the data the high values are removed.

```{r VCFsummary, include=T, message=FALSE}
hist(vcfGewata)
```


```{r VCFtransform}
# Transform VCF
vcfGewata[vcfGewata > 100] <- NA
```

After the removal of the redundant data of the VCF product the bands need to be rescaled. The values of the bands are stored as an integer to save file storage space, but reflectance values have an original scale from 0 to 1.

By producing a rasterbrick we can rescale all bands at once. After this transformation the VCF layer is added to the brick and all layers are named properly


```{r Bandstransform, message=FALSE, include=TRUE, results='hide'}
#Transform Bands
rasterbrick <- brick(GewataB1, GewataB2, GewataB3, GewataB4, GewataB5, GewataB7)
rasterbrick <- calc(rasterbrick, fun=function(x) x / 10000)
rasterbrick <- addLayer(rasterbrick, vcfGewata)
names(rasterbrick) <- c("Band1","Band2", "Band3", "Band4", "Band5", "Band7", "VCF")

```

Now all values will be converted to a dataframe. This will make the plotting possible. First all values are extracted from the rasterbrick. secondly the values are convertet to a data frame. The first 10 rows are shown below.

```{r RastertoData, message=FALSE, include=TRUE, results='hide'}
#Create dataframe
rasterbrickData <- getValues(rasterbrick)
rasterbrickData <- as.data.frame(rasterbrickData, na.rm=T)
```

```{r headRastertoData, message=FALSE, include=TRUE}
head(rasterbrickData)
```

After this conversion the plot can be made.

```{r Plotcorrelation, message=FALSE, include=TRUE}
# Plotting correlation between VCF and different LandSat bands
plot(VCF ~ Band1, data = rasterbrickData, pch = ".", col = "orange", 
     main="Correlation between VCF and LandSat bands", 
     xlab="Reflectance of LandSat bands", ylab="VCF [%]", 
     xlim=c(0,0.5), ylim=c(0,100))

points(VCF ~ Band5, data = rasterbrickData, pch = ".", col = "dark green")
points(VCF ~ Band7, data = rasterbrickData, pch = ".", col = "light blue")
legend(0.37, 100, box.col='white', legend=c("Band 1", "Band 5", "Band 7"),
       fill=c("orange", "dark green", "light blue"), bg="white")
```

## Create a model and show a summary of the model object
In this step to models are created. The first one creates a regression formula from the correlation between LandSat bands 1,5 and 7. The second one takes all bands except the thermal band into account.

```{r Models, message=FALSE, include=TRUE, results='hide'}
# Creating linear regression model
model157 <- lm(VCF ~ Band1+Band5+Band7, rasterbrickData, na.action=na.omit)
modelAll <- lm(VCF ~ Band1+Band2+Band3+Band4+Band5+Band7, rasterbrickData, na.action=na.omit)
```

The summary of the model gives information about the parameters in the formula. From this summary the most important predictors can be derived.

In the summary of the first models all bands are importand predictors (the 'Pr' column under Coefficients). But, if we look at the summary of the second model we can clearly see that for this model the 7th band is not an important predictor.

```{r SummaryModel, message=FALSE, include=TRUE}
summary(model157)
summary(modelAll)
```

## Plot the predicted tree cover raster and compare it with the original

The created models can be used to predict the treecover. For this prediction the 'predict' function is called. A name is assigned to the new raster object and again all possible redundant data is removed (e.g pixels with less than 0 percent tree cover)

```{r Predict, message=FALSE, include=TRUE, results='hide'}
# Predicting VCF based on linear regression model
VCFpredict <- predict(rasterbrick, model = model157, na.rm=T)
names(VCFpredict) <- "VCF"
VCFpredict[VCFpredict < 0] <- NA
VCFpredict[VCFpredict > 100] <- NA
```

To check the results of the prediction, the predicted VCF Tree cover raster can now be compared to the original master. If the model is working properly the resulting plots should be mostly the same.

```{r compareVCF, message=FALSE, include=TRUE, strip.white=TRUE, fig.width=14}
# Comparing predicted and original VCF
colorPal <- rev(colorRampPalette(c("darkgreen","yellow","brown"))(20)) # Create color palette
plot1<-spplot(rasterbrick$VCF, main="Original VCF", col.regions = colorPal)
plot2<-spplot(VCFpredict,main="Predicted VCF", col.regions = colorPal)
print(plot1, position = c(0,0,.5,1),more = T)
print(plot2, position = c(.5,0,1,1),more = T)
```

## Calculate RMSE between predicted and original rasters

And indication of the error of the prediction of the VCF is given by the RMSE (Root Mean Square Error), more information can be found on [wikipedia](http://en.wikipedia.org/wiki/Root-mean-square_deviation). A function is made which returns the RMSE between two raster inputs; the original and predicted raster.

```{r Calculating RMSE, message=FALSE, include=TRUE}
source("R/calculate.RMSE.R")
RMSEpredict <- calculate.RMSE(rasterbrick$VCF, VCFpredict)
paste("The RMSE is", format(round(RMSEpredict,2), nsmall = 2))
```

The function that is used for the RMSE calculation is shown below.

```{r Function RMSE, message=FALSE, include=TRUE}
calculate.RMSE <- function (x,y){
  # x is Original VCF layer, y is the Predicted VCF layer
  minus<-x-y
  square<-minus^2
  mean<-cellStats(square, 'mean')
  RMSE <- sqrt(mean)
  return(RMSE)
}
```

## Calculate RMSE per class
To investigate whether each class can be predicted equally good by the same model, the RMSE between the predicted and original values of the trainingPolygons is calculated per class. First the VCF of the trainingspolygons have to be extracted from both the original and the predicted raster and then the classes are linked to the right polygon. This is done in a seperate function.

```{r Function mask, message=FALSE, include=TRUE}
mask.trainingdata <- function(x,y){
  # x is a VCF raster, y stands for the training areas
  trainingareas <- mask(x,y)
  y@data$Code <- as.numeric(y@data$Class) #select the variable code from the trainingareas
  classes <- rasterize(trainingPoly, x, field = 'Class') #convert the date to a raster
  names(classes) <- 'Class' 
  trainingclass <- addLayer(trainingareas, classes) #add the new layer
  return(trainingclass)
}
```

```{r retreiving trainingsPolygon data, message=FALSE, include=TRUE, results='hide'}
source("R/mask.trainingdata.R")
trainingOri <- mask.trainingdata(rasterbrick$VCF, trainingPoly)
trainingPred <- mask.trainingdata(VCFpredict, trainingPoly)
```

Then the RMSE's are calculated per classes. Again a seperate function is used to select the classes.

```{r Function class, message=FALSE, include=TRUE}
calculate.RMSE.class <- function(r1,r2,class){
  # r1 is the original VCF data in the training area, r2 is the Predicted VCF. 
  r1sub <- r1$VCF #select the VCF column
  r1sub[r1$Class != class] <- NA #all values in unwanted classes are set to NA
  r2sub <- r2$VCF
  r2sub[r2$Class != class] <- NA
  RMSEsub <- calculate.RMSE(r1sub,r2sub)
  return(RMSEsub)
}
```

```{r Calculating RMSE per class, message=FALSE, include=TRUE}
source("R/calculate.RMSE.class.R")
## Calculating RMSE per class
RMSE_crop <- calculate.RMSE.class(trainingOri, trainingPred,1)
RMSE_for <- calculate.RMSE.class(trainingOri, trainingPred,2)
RMSE_wet <- calculate.RMSE.class(trainingOri, trainingPred,3)
```

And printed to compare them:
```{r Print RMSE per class, message=FALSE, include=TRUE}
paste("RMSE of cropland, forest and wetland are",format(round(RMSE_crop, 2), nsmall = 2),",",
      format(round(RMSE_for, 2), nsmall = 2),"and",format(round(RMSE_wet, 2), nsmall = 2),"respectively.")
```

Apparently some classes are predicted better then others. This can be caused because there are not enough training areas to define a class or classes can overlap each others.